---
title: "Castellano_CS636_Lab03"
output:
  html_notebook: default
  pdf_document: default
  author: "Andres Castellano"
  date: "2/15/2020" 
---


# Question #1

Please read the train.csv file into R and store the data in a variable called “X”.

```{r}
library(readr)
X <- read_csv('C:/Users/Castellano/Documents/Spring2020/CS636/Home Depot/train.csv/train.csv')
```

# Question #2

Write a function, called “distinct_relevance”, to count how many distinct values are in the column “relevance”? So when we call the function, it returns the desired results:
distinct_relevance (vect = X$relevance);

```{r}
distinct_relevance <- function(vector) {
  elementos <- c(rep(0,length(vector)))
  for (i in 1:length(vector)) {
    boo <- vector[i] %in% elementos
   if( boo == FALSE) {
    #print('Not in')
     elementos[i] <- vector[i]
   } else {
      next
    }
  }
  elementos <- elementos[!elementos %in% 0]
  return(elementos)
}

relevant <- distinct_relevance(X$relevance)
print(relevant) 

```
3, Write a function, called “count”, to count the number of appearances of a value, e.g. 3, in the column “relevance”, so when we call the function, it returns the desired results:
count(vect = X$relevance, value=3);
(For Q2 and Q3, please do not use existing R packages or functions.)
```{r}
count <- function(vector, value) {
  cuenta <- 0
  for (i in 1:length(vector)) {
   if( vector[i] == value) {
     cuenta <- cuenta + 1 
   }
  }
  return(cuenta)
}

count(X$relevance,3.00)
```


# Question #4 

Compare the results with R function: table()

```{r}
system.time(relevance_values <- distinct_relevance(X$relevance))
relevance_counts <- count(X$relevance,3.00)
system.time(table(X$relevance))
```
Hmmmmmm, the table() function is A LOT more efficient. Wondering why.

5. Pi can be computed by adding the following terms (http://en.wikipedia.org/wiki/Pi): How many terms does it take to get the first 3 digits to be correct, 3.14? Write an R function getPi(N) to compute it, where N specifies the first N digits to be correct, and returns #terms.


```{r}
getPi <- function(N){
  'This function only works for N < 10.
  If a higher precision is desired, this function can be modified as
  an implementation of a while loop instead.'
  old_pie <- 4
  for (k in 1:10^N){
    new_pie <- old_pie + 4*((-1)^k)/(2*k+1)
     #print(new_pie)
  if ( abs(new_pie-old_pie) < 1*10^(-(N-1)  )) {
    k <- paste(toString(k), "Iterations Required")
    new_pie <- round(new_pie, digits = N + 1) 
    Y <- list(k, new_pie) 
    break
  }
    old_pie <- new_pie
  }
  
  return(Y)
}
N <- getPi(3)
print(N)
class(N)
```














































