---
title: "Castellano_CS636_Lab04"
output:
  html_notebook: default
  pdf_document: default
  author: "Andres Castellano"
  date: "2/15/2020" 
---

# 1. **IRIS**

## (1) How many features are numerical data? And how many are categorical data

```{r}
str(iris[0,])
```
There are 5 features, 4 numerical and 1 categorical.

## (2) Histogram 
Make histogram for the numerical features, to see how they distribute

```{r}
par(mfrow=c(2,2))
hist(iris$Petal.Length,main='Histogram of Petal Lenght',xlab='Petal Length (mm)',col = '#9933FF')
hist(iris$Petal.Width,main='Histogram of Petal Width',xlab='Petal Width (mm)', col = '#9933FF')
hist(iris$Sepal.Length,main='Histogram of Sepal Lenght',xlab='Sepal Length (mm)', col = '#CC99FF')
hist(iris$Petal.Length,main='Histogram of Sepal Width',xlab='Sepal Width (mm)', col = '#CC99FF')
```
## (3) Table 
Make table for the categorical features, to see how they distribute
```{r}
table(iris$Species)
barplot(table(iris$Species), main = 'Distribution of Species', xlab = 'Species', ylab = 'Frequency',space = .05, col = '#99ff33', ylim = c(0,70) ) 


```


# 2.  **Rivers**

The data set rivers contains the lengths (in miles) of 141 major rivers in North America.

## (1) What proportion are less than 500 miles long?

```{r}
t = length(rivers[rivers < 500])
p = t/length(rivers) 
p
```

## (2) What proportion are less than the mean length?

```{r}
t = length(rivers[rivers < mean(rivers)])
p = t/length(rivers)
p
```

## (3) What is the 0.75 quantile?

```{r}
third = quantile(rivers)
third
```

The 0.75 quantile is 680miles.

# 3. **pi2000**

Fit a density estimate to the data set pi2000 (UsingR). Compare with the appropriate histogram. Why might you want to add an argument like breaks =0:10-0.5 to hist()?

```{r}

data(pi2000)
par(mfrow=c(1,2))
x <- pi2000
h <- hist(x, breaks = 0:10-0.5, col="light green", xlab='Digits', main='Histogram of Pi Digits',prob=T)
xfit <- seq(min(x),max(x),length=9)
yfit <- dunif(xfit,0,9)
yfit <- yfit*.8
lines(xfit,yfit,col='blue',lwd=2)

h <- hist(x, breaks = 0:10, col="light green", xlab='Digits', main='Histogram of Pi Digits',prob=T)
xfit <- seq(min(x),max(x),length=9)
yfit <- dunif(xfit,0,9)
yfit <- yfit*.8
lines(xfit,yfit,col='blue',lwd=2)
```

Using breaks =  0:10-0.5 helps display the histogram more accurately since using breaks = 10 for example, causes the histogram to be plotted out of proportion.

# 4. **MASS**

The data set DDT (MASS) contains independent measurements of the pesticide
DDT on kale. Make a histogram and a boxplot of the data. From these, estimate the mean and standard deviation. Check your answers with the appropriate functions.

```{r}

hist(DDT,breaks=5,col="orange", main="Histogram of DDT mass in Kale")
boxplot(DDT,col="orange",main="BoxPlot of DDT mass in Kale")
```
### **MASS Estimates**

$mean \approx 3.6$

$sd \approx 0.4$ 

### **MASS Actuals**

```{r}
str(DDT)
sprintf('Mean mass is %s ', mean(DDT))
sprintf('Standard deviation is %.2f ', sd(DDT))
```


# 5. **Two Graphics**

It can be illuminating to view two different graphics of the same data set at once.
A simple way to stack graphics is to specify that a figure will contain two graphics by using the command

>par(mfrow=c(2,1) 

>hist(x)
 
>boxplot(x, horizontal=TRUE)


will produce stacked graphics. (The graphics device will remain divided until you change it back with a command such as par (mfrow=c(1, 1)) or close the device.) For the data set lawsuits (UsingR), make stacked graphics of lawsuits and log (lawsuits). Could you have guessed where the middle 50% of the data would have been without the help of the boxplot?

```{r}
par(mfrow=c(2,2)) 
x <- (lawsuits)
boxplot(x, col="#0099ff", main='Boxplot', xlab='Lawsuits')
boxplot(log(x), col='#0099ff', horizontal=TRUE, main='Boxplot', 
xlab='log(Lawsuits)')
hist(x, col="#0099ff", main='Histogram', xlab='Lawsuits')
hist(log(x), col='#0099ff', main='Histogram', 
xlab='log(Lawsuits)')
```
Determining the median of the data would have been impossible without the aid of box plot of log(lawsuits).


# 6. **Sex, Age, and Smoking** 
Let sex = c(1,1,1,1,2,2,2,2,2,2); smoking = c(1,0,1,0,1,0,0,0,1,1); age=c(31:40) in R. A data frame is constructed as zz = data.frame(sex, smoking, age). Give the results of following R commands:
1)	apply(zz[-1,], 2, min)

```{r}
sex <- c(1,1,1,1,2,2,2,2,2,2)
smoking <- c(1,0,1,0,1,0,0,0,1,1)
age <- c(31:40)
zz <- data.frame(sex, smoking, age)
zz
```
```{r}
apply(zz[-1,], 2, min)
zz[zz[,3]>35,]
zz[order(zz["smoking"], zz["age"]),  ]
subset(zz, zz["sex"]==1)
tapply(zz$age, zz$smoking, max)
apply(zz[,-3], 1, function(x){ sum(x) })

```


