---
title: "R Notebook"
output: html_notebook
---

# CS636 Homework 2

Please submit the code together with the running results of the testing commands.

Please do not use existing R packages and functions like sort(), order() and so on.

## 1) Piecewise Function
Write a function, F(x), which takes x as the input parameter. It calculates and prints the value of the following mathematical function.


```{r}
F <- function(x){
  if (x < 1){
    f <- 5/(x-1)**2}
  else if ( x == 1){
    f <- 2}
  else {
    f <- 5/(x-1)**3}
  return(f)
  print(f)
  }
```

Testing commands:	F(1);	F(10);	F(0.3);

```{r}
F(0.3)
```

# 2) Fibonacci
The Fibonacci sequence 1, 1, 2, 3, 5, 8, 13, 21…… starts with two 1s, and each term afterwards is the sum of its two predecessors. Please write a function, Fib(n), which takes n as the input parameter. It will return the n-th number in the Fibonacci sequence.

```{r}
Fib <-  function(n){
  ni <- 1
  nj <- 1
  if (n <= 2){
    fib <- ni
    return(fib)}
  else if ( n > 2){
    for (i in seq(3,n)){
      fib <- ni + nj
      ni <- nj
      nj <- fib}
    return(fib)
    }
}
```



Testing commands:	Fib(1);	Fib(2);	Fib(100);
 
```{r}
Fib(8) #Tested with Fib(1). Fib(2), Fib(8)
```

# 3 Merge Algorithm
The merge operation plays an important role in merge sort algorithm. Suppose you have two sorted sequences S1 and S2, merge operation will combine these two sequences into a single ordered sequence. Please write a function, Merge(S1, S2), which accepts two ordered vectors S1 and S2 as parameters. It will return a single ordered sequence. For example,
S1 = c(1,3,5,7);
S2=c(2,4,6,10);
Merge(S1, S2) will return c(1,2,3,4,5,6,7,10)

```{r}
Merge <- function(s1 , s2){
  fin <- c()
  size_1 <- length(s1)
  size_2 <- length(s2)
  
  
  i <- 1
  j <- 1
  
  while (i < size_1 && j < size_2){
    if (s1[i] < s2[j]){
      fin <- append(fin, s1[i])
      i <- i +1}
    else {
      fin <- append(fin, s2[j])
      j <- j +1}
  }
  fin <- append(fin,c(s1[i:size_1],s2[j:size_2]))
}
```


Testing commands:	Merge(seq(1, 50, by=3), seq(2, 30, by=2))

```{r}
res <- Merge(seq(1, 50, by=3), seq(2, 30, by=2))
print(res)
```


# 4) Quick Sort Algorithm
One of the most important algorithms is the quick sort, which is based on the quick sort partition. Here we implement a simple version of the partition function. Please write a function, Partition(pivot, vect), which takes two parameters. The function partitions the sequence, vect, into two parts (part1 <= pivot; part2 > pivot) based on the pivot. For example,
Pivot = 6;
Vect = c(1, 5, 3, 7, 9, 6, 4, 2, 10, 8);
List = Partition(Pivot, Vect);
List[[1]] is c(1,5,3,4,2, 6) and List[[2]] is c(7, 9, 10, 8).

Note that Partition returns a list.

```{r}
Partition <- function(Pivot, Vect){
  first <- c()
  last <- c()
  for (i in seq(1,length(Vect))){
    if (Vect[i] <= Pivot){
      first<-append(first,Vect[i])}
    else {last<-append(last,Vect[i])}}
  res <- list(first,last)
  return(res)}
```


Testing commands:	Partition(50, sample(1:100, 100, replace=F) 

```{r}
Partition(50, sample(1:100, 100, replace=FALSE))
```


